Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-12-02 src/tex/ms.tex

% Define document class
\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
% \documentclass[linenumbers]{aastex631}

@@

The two-infall scenario was first proposed by \citet{chiappini_chemical_1997} to explain the origin of the $\alpha$-bimodality in the MW. The basic premise is that the rate of gas accretion onto the disk occurs in two bursts. The relatively low infall rate between the two bursts causes star formation to slow down, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars. 
The infall timescale for the low-$\alpha$ disk can vary with radius to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the Solar neighborhood abundance data available at the time.

Subsequent studies refined the two-infall scenario to reproduce abundance data across the disk \citep[e.g.,][]{chiappini_abundance_2001,chiappini_oxygen_2003}, as well as the local surface densities of stars and gas and the local star formation and supernova rates \citep[e.g.,][]{romano_mass_2000,spitoni_galactic_2020,spitoni_remind_2024}. Others have explored the SN Ia delay-time distribution \citep{matteucci_effect_2009,palicio_analytic_2023}, galactic fountains \citep{spitoni_effects_2009}, radial gas flows \citep{spitoni_effects_2011,palla_chemical_2020}, variations in the star formation efficiency \citep{spitoni_effects_2011,palla_chemical_2020}, radial stellar migration \citep{spitoni_effect_2015,palla_mgfe_2022}, azimuthal abundance variations \citep{spitoni_2d_2019}, and pre-enriched gas infall \citep{palla_chemical_2020,spitoni_remind_2024}. \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation (AMR), which could be interpreted as evidence supporting the two-infall scenario, but \citet{plotnikova_chemical_2024} argued that such patterns can arise due to selection effects. \citet{spitoni_beyond_2023} and \citet{palla_mapping_2024} proposed a third gas accretion event in the last $\sim3$ Gyr to match the star formation history inferred from {\it Gaia} \citep{ruiz-lara_recurrent_2020} and to explain the recent abundance evolution (or lack thereof) in the Solar neighborhood. Most recently, \citet{hegedus_reconstructing_2025} constrained the infall timescales using chemical abundances for $\sim400,000$ stars from the 19th data release of the Sloan Digital Sky Survey \citep[SDSS;][]{sdss_collaboration_nineteenth_2025}.

@@

@@

We run multi-zone GCE models using the Versatile Integrator for Chemical Evolution \citep[{\tt VICE};][]{johnson_impact_2020}. The basic format of our models follows \citet{johnson_stellar_2021} and \citet{dubay_galactic_2024}. We set up a disk spanning $0\leq R_{\rm gal}<20\kpc$ that is divided into concentric rings of width $\delta R_{\rm gal}=100\,{\rm pc}$. We use a time-step of $\Delta t=10\,{\rm Myr}$, $n=8$ stellar populations per time-step per ring, and we run our models to a final time of $t_{\rm final}=13.2\,{\rm Gyr}$. Within each ring, chemical evolution proceeds according to a conventional one-zone GCE model with instantaneous mixing and continuous recycling. Stellar populations migrate between zones, allowing the long-lived progenitors of SNe Ia to enrich areas of the Galaxy outside of their birth zones, which couples the enrichment in nearby rings. We inhibit star formation past $R_{\rm gal}>15.5\kpc$, so stars in the outer 4.5 kpc of the model disk represent a purely migrated population.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the $e$-folding timescale of the star formation history, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. Motivated by Equation \ref{eq:equilibrium}, we adopt a different outflow mass-loading factor at the Solar radius $\eta_\odot\equiv\eta(R=R_\odot)$ for each of the yield sets in Table \ref{tab:yields}. 
Measurements of gas-phase \citep[e.g.,][]{mendez-delgado_gradients_2022} and stellar (Figure \ref{fig:yield-outflow}) abundances indicate that the Solar neighborhood ISM is presently close to Solar metallicity, so we adjust $\eta_\odot$ to ensure that $Z_{\rm O,eq}\approx Z_{\rm O,\odot}$.
Table \ref{tab:models} reports the value of $\eta_\odot$ in each of our models.

@@

>>>>

% Define document class
\documentclass[trackchanges,twocolumn,twocolappendix,linenumbers]{aastex631}

@@

The two-infall scenario was first proposed by \citet{chiappini_chemical_1997} to explain the origin of the $\alpha$-bimodality in the MW. The basic premise is that the rate of gas accretion onto the disk occurs in two bursts. The relatively low infall rate between the two bursts causes star formation to slow down, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars. 
% The infall timescale for the low-$\alpha$ disk can vary with radius to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
% The initial model of 
\citet{chiappini_chemical_1997} successfully reproduced the Solar neighborhood abundance data available at the time, and
subsequent studies refined the two-infall scenario to reproduce abundance data across the disk \citep[e.g.,][]{chiappini_abundance_2001,chiappini_oxygen_2003}, as well as the local surface densities of stars and gas and the local star formation and supernova rates \citep[e.g.,][]{romano_mass_2000,spitoni_galactic_2020,spitoni_remind_2024}. Others have explored the SN Ia delay-time distribution \citep{matteucci_effect_2009,palicio_analytic_2023}, galactic fountains \citep{spitoni_effects_2009}, radial gas flows \citep{spitoni_effects_2011,palla_chemical_2020}, variations in the star formation efficiency \citep{spitoni_effects_2011,palla_chemical_2020}, radial stellar migration \citep{spitoni_effect_2015,palla_mgfe_2022}, azimuthal abundance variations \citep{spitoni_2d_2019}, and pre-enriched gas infall \citep{palla_chemical_2020,spitoni_remind_2024}. \citet{spitoni_beyond_2023} and \citet{palla_mapping_2024} proposed a third gas accretion event in the last $\sim3$ Gyr to match the star formation history inferred from {\it Gaia} \citep{ruiz-lara_recurrent_2020} and to explain the recent abundance evolution (or lack thereof) in the Solar neighborhood. Most recently, \citet{hegedus_reconstructing_2025} constrained the infall timescales using chemical abundances for $\sim400,000$ stars from the 19th data release of the Sloan Digital Sky Survey \citep[SDSS;][]{sdss_collaboration_nineteenth_2025}.

@@

\todo{Integrate into previous para:} \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation (AMR), which could be interpreted as evidence supporting the two-infall scenario, but \citet{plotnikova_chemical_2024} argued that such patterns can arise due to selection effects.
\todo{Also: multiple sequences in Anders et al. (2023), Xiang \& Rix (2022), Sahlholdt et al. (2022).}

@@

We run multi-zone GCE models using the Versatile Integrator for Chemical Evolution \citep[{\tt VICE};][]{johnson_impact_2020}. The basic format of our models follows \citet{johnson_stellar_2021} and \citet{dubay_galactic_2024}. We set up a disk spanning $0\leq R_{\rm gal}<20\kpc$ that is divided into concentric rings of width $\delta R_{\rm gal}=100\,{\rm pc}$. We use a time-step of $\Delta t=10\,{\rm Myr}$, $n=8$ stellar populations per time-step per ring, and we run our models to a final time of $t_{\rm final}=13.2\,{\rm Gyr}$. \added{We adopt a \citet{kroupa_variation_2001} initial mass function (IMF).} Within each ring, chemical evolution proceeds according to a conventional one-zone GCE model with instantaneous mixing and continuous recycling. Stellar populations migrate between zones, allowing the long-lived progenitors of SNe Ia to enrich areas of the Galaxy outside of their birth zones, which couples the enrichment in nearby rings. We inhibit star formation past $R_{\rm gal}>15.5\kpc$, so stars in the outer 4.5 kpc of the model disk represent a purely migrated population.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter \added{for a \citet{kroupa_variation_2001} IMF}, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the $e$-folding timescale of the star formation history, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. Motivated by Equation \ref{eq:equilibrium}, we adopt a different outflow mass-loading factor at the Solar radius $\eta_\odot\equiv\eta(R=R_\odot)$ for each of the yield sets in Table \ref{tab:yields}. 
Measurements of gas-phase \citep[e.g.,][]{mendez-delgado_gradients_2022} and stellar (Figure \ref{fig:yield-outflow}) abundances indicate that the Solar neighborhood ISM is presently close to Solar metallicity, so we adjust $\eta_\odot$ to ensure that $Z_{\rm O,eq}\approx Z_{\rm O,\odot}$.
Table \ref{tab:models} reports the value of $\eta_\odot$ in each of our models.

@@

\todo{Contrast with: multiple sequences in Anders et al. (2023), Xiang \& Rix (2022), Sahlholdt et al. (2022).}
